### 2019-01-04 manuscript/template.tex
updated/fixed cliff-walking figure.

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript,palatino}
\usepackage{graphicx}

\title{something something about anxiety}

\author{
Samuel Zorowitz \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{zorowitz@princeton.edu} \\
\And
Nathaniel Daw \\
Princeton Neuroscience Institute\\
Princeton University\\
Princeton, NJ 08540 \\
\texttt{ndaw@princeton.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
To be filled in
\end{abstract}

\keywords{
reinforcement learning; avoidance; anxiety
}

\startmain % to start the main 1-4 pages of the submission.

\section{Submission of papers to RLDM}

RLDM requires electronic submissions.  This year's electronic
submission site is
\begin{center}
   https://cmt3.research.microsoft.com/RLDM2017/
\end{center}

Please read the instructions below, and follow them faithfully. Note
that there is also a template \verb+rldm.rtf+ for Microsoft Word,
which is available from the website below.
\subsection{Style}

Papers to be submitted to RLDM must be prepared according to the
instructions presented here. Papers consist of a \emph{title}, which
has a maximum of 100 characters, an \emph{abstract}, which is a
maximum of 2000 characters, up to five key words, and an
\emph{extended abstract}, which starts on the second page, and can be
between one and four pages. Figures and references should be included
in the latter.

Authors preferring \LaTeX{} are requested to use the RLDM \LaTeX{}
style files obtainable at the RLDM website at
\begin{center}
   http://www.rldm.org/
\end{center}
The file \verb+rldm.pdf+ contains these instructions and illustrates
the various formatting requirements your RLDM paper must
satisfy. There is a \LaTeX{} style file called \verb+rldmsubmit.sty+,
and a \LaTeX{} file \verb+rldm.tex+, which may be used as a ``shell''
for writing your paper. All you have to do is replace the author,
title, abstract, keywords, acknowledgements and text of the paper with
your own. The file
\verb+rldm.rtf+ is provided as an equivalent shell for Microsoft Word users.

\section{General formatting instructions}
\label{gen_inst}

The paper size for RLDM is ``US Letter'' (rather than ``A4''). Margins
are 1.5cm around all sides. Use 11~point type with a vertical spacing
of 12~points. Palatino is the preferred typeface throughout.
Paragraphs are separated by 1/2~line space, with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 0.6cm space above and below title to rules.

The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

\section{Preparing PostScript or PDF files}

Please prepare PostScript or PDF files with paper size ``US Letter''.
The -t letter option on dvips will produce US Letter files.

\end{document}